---
title: "Week-4 assignment"
author: "Palash Vinay Siddamsettiwar"
date: "November 2, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Synopsis
In this assignment, we learn how to use the data from gapminder package, and perform exploratory data analysis (EDA) on this data though data manipulation packages such as _dplyr_ and data visualizations from packages such as _ggplot2_. We find various parameters around gdp per capita growth over the last 55 years in the world generally and in my country of origin, India, in particular.

##Packages Required

**ggplot2** - This package helps with creation of plots in R
``` {r, warning=F, message=F}
#install.packages("ggplot2")
library("ggplot2")
```

**dplyr** - This package contains a lot of great data manipulation functions (part of tidyverse)
``` {r, warning=F, message=F}
#install.packages("tidyverse")
library("tidyverse")
```

**gapminder** - This package provides gapminder data as dataframes in R
``` {r, warning=F, message=F}
#install.packages("gapminder")
library("gapminder")
```

##Source Code
From the R function '?gapminder_unfiltered' and '?gapminder', we can see that the data contains 3313 observations across 6 variables. The 6 varaibles and their descriptions are: 

1. **country**: Lists all the country names as a factor (187 levels)

2. **continent**: Lists all the continent names as a factor (6 levels)

3. **year**: Year ranging from 1952 to 2007 in increments of 1 year

4. **lifeExp**: Life expectation at birth, in years, for the country

5. **pop**: Population of the country

6. **gdpPercap**: Gross Domestic Product per capita for the country

**Missing Values:** `r colMeans(is.na(gapminder_unfiltered))*100`

**Continents Included:** `r unique(gapminder_unfiltered$continent)[order(unique(gapminder_unfiltered$continent))]`

**Number of Countries:** `r length(unique(gapminder_unfiltered$country)) `

**Years Measured:** `r  paste0(min(unique(gapminder_unfiltered$year)), "-", max(unique(gapminder_unfiltered$year))) `

##Exploratory Data Analysis
**1. For the year 2007, what is the distribution of GDP per capita across all countries?**
``` {r, warning=FALSE}
boxplot(x = filter(gapminder_unfiltered, year==2007)$gdpPercap)
ggplot(data = filter(gapminder_unfiltered, year==2007)) + 
  geom_point(mapping = aes(x = country, y = gdpPercap))+
  geom_text(label=gapminder_unfiltered$country)+
  theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank())
```

**2. For the year 2007, how do the distributions differ across the different continents?**
``` {r }
ggplot(data = filter(gapminder_unfiltered, year==2007)) + 
  geom_point(mapping = aes(x = country, y = gdpPercap)) + 
  facet_wrap(~ continent, nrow = 2)+
    theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank())
```

**3. For the year 2007, what are the top 10 countries with the largest GDP per capita?**
``` {r }
arrange(filter(gapminder_unfiltered, year==2007)[c(1,6)], desc(gdpPercap))[c(1:10),]
```

**4. Plot the GDP per capita for your country of origin for all years available.**
``` {r }
country_of_origin="India"
ggplot(data = filter(gapminder_unfiltered, country==country_of_origin), aes(x = year, y=gdpPercap)) + 
  geom_bar(stat='identity')
```

**5. What was the percent growth (or decline) in GDP per capita in 2007?**
``` {r }
a <- filter(gapminder_unfiltered, year %in% c(2007,2002))
a <- group_by(a, year)
b<- summarise(a, avg_growth_gdpPercap = mean(gdpPercap, na.rm = TRUE))
paste0("Percent growth (or decline) in GDP per capita in 2007 for the world was ", round(b[2,2]/b[1,2], 2), "%")
```

**6. What has been the historical growth (or decline) in GDP per capita for your country?**
``` {r, warning=F}
country_of_origin="India"
india_data = filter(gapminder_unfiltered, country==country_of_origin)
a <- as.data.frame(cbind(india_data$year, round((india_data$gdpPercap/lag(india_data$gdpPercap,1) - 1)*100, 2)))
names(a) <-  c("Year", "Growth")
a
paste0("Average histoical growth (or decline) in GDP per capita for India is ", round(mean(a[-1,2], na.rm = T), 2), "%")

ggplot(a, aes(Year, Growth)) + geom_line() +
 xlab("Year") + ylab("Growth % in GDP per capita")
```


gapminder_unfiltered %>%
  filter(year==2007) %>%
  ggplot(aes("", gdpPercap))+
  coord_flip()+
  geom_text(aes(label=country), angle=90)+
  coord_flip()
  
  
gapminder_unfiltered %>%
  filter(year==2007) %>%
  